---
title: "Peri-INTACT GFP Integration"
author: "Jonathan Nelson"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: 
    toc: yes
    toc_depth: 3
    toc_float: true
    number_sections: yes
    theme: bootstrap
    df_print: paged
    code_folding: hide
    highlight: pygments
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Packages

```{r load new packages, echo=TRUE, error=FALSE, message=FALSE, warning=FALSE}

if (!require("knitr")) {install.packages("knitr"); require("knitr")}
if (!requireNamespace('BiocManager', quietly = TRUE)) {install.packages('BiocManager'); require("BiocManager")}
if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
if (!require("stringr")) {install.packages("stringr"); require("stringr")}
if (!require("Seurat")) {install.packages("Seurat"); require("Seurat")}
if (!require("sctransform")) {install.packages("sctransform"); require("sctransform")}
if (!require("glmGamPoi")) {BiocManager::install('glmGamPoi'); require("glmGamPoi")}
if (!require("patchwork")) {install.packages("patchwork"); require("patchwork")}
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
if (!require("EnhancedVolcano")) {BiocManager::install('EnhancedVolcano'); require("EnhancedVolcano")}
if (!require("DESeq2")) {BiocManager::install('DESeq2'); require("DESeq2")}
if (!require("tidyverse")) {install.packages("tidyverse"); require("tidyverse")}
if (!require("RColorBrewer")) {install.packages("RColorBrewer"); require("RColorBrewer")}
if (!require("car")) {install.packages("car"); require("car")}
if (!require("openxlsx")) {install.packages("openxlsx"); require("openxlsx")}
if (!require("readxl")) {install.packages("readxl"); require("readxl")}
if (!require("ggrepel")) {install.packages("ggrepel"); require("ggrepel")}
if (!require("gghighlight")) {install.packages("gghighlight"); require("gghighlight")}
if (!require("ggpmisc")) {install.packages("ggpmisc"); require("ggpmisc")}
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
if (!require("here")) {install.packages("here"); require("here")}
```

# Loading Filtered Seurat Object
These files are post-DoubletFinder.

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

One <- readRDS(here("Outputs", "951_260724.rds"))
Two <- readRDS(here("Outputs", "3246_260724.rds"))

One <- AddMetaData(object = One, metadata = "951", col.name = "Sample") #Change
Two <- AddMetaData(object = Two, metadata = "3246", col.name = "Sample") #Change

```

# Nuclei Entering the Analysis Pipeline
The following shows the number of nuclei post-QC and post-DoubletFinder. Numbers in `# (10X Count)` are the starting numbers for reference.  
Values appear in the Environment when running the `rmd`.

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}
One.nuc <- nrow(One@meta.data) # 10040 (10X Count)
One.nuc
Two.nuc <- nrow(Two@meta.data) # 11725 (10X Count)
Two.nuc
```

# Merging Replicates

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

GFPcombined <- merge(One, y = c(Two), project = "Peri-INTACT GFP")

```

## Nuclei Filtering and Batch Comparison

```{r echo=FALSE, error=FALSE, message=FALSE, warning=FALSE}

head(GFPcombined@meta.data)

VlnPlot(GFPcombined, 
        features = c("nCount_RNA", "nFeature_RNA", "percent.mt"),
        group.by = "Sample")

GFPcombined <- subset(GFPcombined,
                                  nFeature_RNA < 10000 &
                                  nCount_RNA > 1000 &
                                  nCount_RNA < 10000) 

VlnPlot(GFPcombined, 
        features = c("nCount_RNA", "nFeature_RNA", "percent.mt"),
        group.by = "Sample")

VlnPlot(GFPcombined, 
        features = c("nCount_RNA", "nFeature_RNA", "percent.mt"),
        group.by = "Sample", pt.size = 0)


GFPcombined

```


## Data Normalization

```{r echo=TRUE, warning=FALSE, error=FALSE, results = FALSE, message=FALSE, fig.show='hide', fig.align = 'center'}

GFPcombined <- SCTransform(GFPcombined, method = "glmGamPoi", vars.to.regress = c("nCount_RNA"), verbose = TRUE)

```

## Principal Component Analysis

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

GFPcombined <- RunPCA(GFPcombined, verbose = TRUE)
ElbowPlot(GFPcombined, ndims = 50)
```

## Dataset Integration
- Normalize datasets individually by `SCTransform()`, instead of `NormalizeData()` prior to integration.  
- Use 3,000 or more features for analysis downstream of `sctransform`.  
- Run the `PrepSCTIntegration()` function prior to identifying anchors.  
- When running `FindIntegrationAnchors()`, and `IntegrateData()`, set the normalization.method parameter to the value `SCT`.  
- When running sctransform-based workflows, including integration, **do not** run the `ScaleData()` function.

```{r echo=TRUE, warning=FALSE, error=FALSE, results = FALSE, message=FALSE, fig.align = 'center'}

GFPcombined.list <- SplitObject(GFPcombined, split.by = "Sample")

GFPcombined.list <- lapply(X = GFPcombined.list, FUN = SCTransform)

features <- SelectIntegrationFeatures(object.list = GFPcombined.list, nfeatures = 3000)

GFPcombined.list <- PrepSCTIntegration(object.list = GFPcombined.list, anchor.features = features)

anchors <- FindIntegrationAnchors(object.list = GFPcombined.list,
                                  reduction = "rpca",
                                  normalization.method = "SCT",
                                  anchor.features = features)

GFPcombined <- IntegrateData(anchorset = anchors,
                          normalization.method = "SCT")

rm(anchors, One, Two)

```

## UMAP Visualization and Tables

```{r echo=TRUE, warning=FALSE, error=FALSE, results = FALSE, message=FALSE, fig.align = 'center'}

DefaultAssay(GFPcombined) <- "integrated"

GFPcombined <- RunPCA(GFPcombined, verbose = FALSE)
GFPcombined <- RunUMAP(GFPcombined, dims = 1:50)
 
GFPcombined <- FindNeighbors(GFPcombined, reduction = "pca", dims = 1:50)
GFPcombined <- FindClusters(GFPcombined, resolution = 3.0)

DimPlot(GFPcombined, reduction = "umap", label = TRUE)

DimPlot(GFPcombined, reduction = "umap", group.by = "Sample")

DimPlot(GFPcombined, reduction = "umap", split.by = "Sample")


f1 <- DimPlot(GFPcombined,
              reduction = "umap",
              label = TRUE,
              pt.size = 0,
              label.size = 3) +
              NoLegend()


f3 <- VlnPlot(GFPcombined, 
        features = c("nCount_RNA", "nFeature_RNA", "percent.mt"),
        fill.by = "ident",
        stack = TRUE, 
        flip = TRUE,
        pt.size = 0) +
 theme(legend.position = "none", 
        plot.title = element_text(hjust = 0.5), 
        axis.text.x = element_text(angle = 0, hjust = .5),
        axis.title.x = element_blank()) +
  stat_summary(fun = median,
        geom = "crossbar",
        width = 0.3,
        size = 0.2,
        position = position_dodge(width = 0.5))

f1 + f3

FeaturePlot(GFPcombined,
             features = c("nCount_RNA", "nFeature_RNA", "percent.mt"),
             cols = c("lightgrey", "royal blue"),
             ncol = 3)

table(GFPcombined$Sample)

table(Idents(GFPcombined))

table(Idents(GFPcombined), GFPcombined$Sample)

```
## Save Intermediate Objects

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

saveRDS(GFPcombined, here("Outputs", "GFPcombined.072624.rds"))

GFPcombined <- readRDS(here("Outputs", "GFPcombined.072624.rds"))

```

# QC the cell populations

## Filter for sample-biased clusters and non-Pdgfrb cells

```{r echo=TRUE, warning=FALSE, error=FALSE, results = FALSE, message=FALSE, fig.align = 'center', fig.width=16}

table(Idents(GFPcombined), GFPcombined$Sample)

DefaultAssay(GFPcombined) <- "SCT"

markers.to.plot1 <- c("Lrp2",         # PT
                      "Slc5a12",      # PT-S1
                      "Slc13a3",      # PT-S2
                      "Slc16a9",      # PT-S3
                      "Havcr1",       # Injured PT
                      "Epha7",        # dTL
                      "Cryab",        # dTL
                      "Slc12a1",      # TAL
                      "Umod",         # TAL, DCT1
                      "Egf",          # TAL, DCT1,
                      "Cldn10",       # TAL
                      "Cldn16",       # TAL
                      "Nos1",         # MD
                      "Slc12a3",      # DCT
                      "Pvalb",        # DCT1
                      "Slc8a1",       # DCT2, CNT
                      "Aqp2",         # PC
                      "Slc4a1",       # IC-A
                      "Slc26a4",      # IC-B
                      "Nphs1",        # Podo
                      "Flt1",         # Endo
                      "Emcn",         # Glom Endo
                      "Kdr",          # Capillary Endo
                      "Pdgfrb",       # Perivascular
                      "Pdgfra",       # Fib
                      "Piezo2",       # Mesangial
                      "Acta2",        # Mural
                      "Ptprc",        # Immune
                      "Top2a"         # Proliferation
                      )

DimPlot(GFPcombined, reduction = "umap", label = TRUE)

DotPlot(GFPcombined,
        features = markers.to.plot1,
        dot.scale = 8,
        dot.min = 0,
        scale.max = 100,
        scale.min = 0,
        col.min = -2.5,
        col.max = 2.5)+
  coord_flip()

VlnPlot(GFPcombined, "nFeature_RNA")

```

## Filtering non-Pdgfrb Clusters

```{r echo=TRUE, warning=FALSE, error=FALSE, results = FALSE, message=FALSE, fig.align = 'center'}

SO <- subset(GFPcombined, idents = c("24", "28", "38", "39", "41", "42", "45", "46", "47", "48", "50"), invert = TRUE)

DimPlot(SO, reduction = "umap", label = TRUE)

#rm(GFPcombined)

table(Idents(SO), SO$Sample)

```

## Reclustering the dataset without doublets

```{r echo=TRUE, warning=FALSE, error=FALSE, results = FALSE, message=FALSE, fig.align = 'center', fig.width=12}

SO <- SCTransform(object = SO)
SO <- RunPCA(object = SO)
SO <- FindNeighbors(object = SO, dims = 1:40)
SO <- FindClusters(object = SO, resolution = 3)
SO <- RunUMAP(object = SO, dims = 1:40)

DimPlot(object = SO, reduction = "umap", label = TRUE, repel = T)

table(Idents(SO), SO$Sample)

DotPlot(SO,
        features = markers.to.plot1,
        dot.scale = 8,
        dot.min = 0,
        scale.max = 100,
        scale.min = 0,
        col.min = -2.5,
        col.max = 2.5)+
  coord_flip()

VlnPlot(SO, "nFeature_RNA")

```


## Annotating the dataset

```{r echo=TRUE, warning=FALSE, error=FALSE, results = FALSE, message=FALSE, fig.align = 'center'}


SO <- SO
SO@meta.data <- SO@meta.data %>% 
  mutate(class.Merge = dplyr::case_when(
    seurat_clusters == 0  ~ "Fibroblasts",
    seurat_clusters == 1  ~ "Fibroblasts",
    seurat_clusters == 2  ~ "Fibroblasts",
    seurat_clusters == 3  ~ "Fibroblasts",
    seurat_clusters == 4  ~ "Fibroblasts",
    seurat_clusters == 5  ~ "Fibroblasts",
    seurat_clusters == 6  ~ "Fibroblasts",
    seurat_clusters == 7  ~ "Fibroblasts",
    seurat_clusters == 8  ~ "Fibroblasts",
    seurat_clusters == 9  ~ "Fibroblasts",
    seurat_clusters == 10 ~ "Fibroblasts",
    seurat_clusters == 11 ~ "Fibroblasts",
    seurat_clusters == 12 ~ "Fibroblasts",
    seurat_clusters == 13 ~ "Fibroblasts",
    seurat_clusters == 14 ~ "Fibroblasts",
    seurat_clusters == 15 ~ "Fibroblasts",
    seurat_clusters == 16 ~ "Pericytes",
    seurat_clusters == 17 ~ "Fibroblasts",
    seurat_clusters == 18 ~ "Fibroblasts",
    seurat_clusters == 19 ~ "Fibroblasts",
    seurat_clusters == 20 ~ "Fibroblasts",
    seurat_clusters == 21 ~ "Pericytes",
    seurat_clusters == 22 ~ "Fibroblasts",
    seurat_clusters == 23 ~ "Afferent Artery",
    seurat_clusters == 24 ~ "Fibroblasts",
    seurat_clusters == 25 ~ "Fibroblasts",
    seurat_clusters == 26 ~ "Fibroblasts",
    seurat_clusters == 27 ~ "Efferent Artery",
    seurat_clusters == 28 ~ "Fibroblasts",
    seurat_clusters == 29 ~ "Pericytes",
    seurat_clusters == 30 ~ "Pericytes",
    seurat_clusters == 31 ~ "Fibroblasts",
    seurat_clusters == 32 ~ "Medullary Fibroblasts",
    seurat_clusters == 33 ~ "Fibroblasts",
    seurat_clusters == 34 ~ "Medullary Fibroblasts",
    seurat_clusters == 35 ~ "Fibroblasts",
    seurat_clusters == 36 ~ "JGA",
    seurat_clusters == 37 ~ "Fibroblasts",
    seurat_clusters == 38 ~ "Mesangial Cells",
    seurat_clusters == 39 ~ "Mesangial Cells",
    seurat_clusters == 40 ~ "Fibroblasts",
    seurat_clusters == 41 ~ "Efferent Artery",
    seurat_clusters == 42 ~ "Fibroblasts",
    seurat_clusters == 43 ~ "Fibroblasts",
    seurat_clusters == 44 ~ "Proliferating",
    seurat_clusters == 45 ~ "Fibroblasts"
    ))

SO@meta.data$class.Merge <- factor(SO@meta.data$class.Merge , levels = c("Fibroblasts", "Medullary Fibroblasts", "Proliferating", "Mesangial Cells", "Pericytes", "Efferent Artery", "Afferent Artery", "JGA"))


DimPlot(object = SO, reduction = "umap", group.by = "class.Merge", label = TRUE)

```

```{r echo=TRUE, warning=FALSE, error=FALSE, results = FALSE, message=FALSE, fig.align = 'center'}


SO3 <- SO

DefaultAssay(SO3) <- "RNA"
SO3 <- NormalizeData(object = SO3)
SO3 <- ScaleData(object = SO3)

markers.to.plot1 <- c("Pdgfrb",
                      "Pdgfra",
                      "Cryab",
                      "Top2a",
                      "Gata3",
                      "Cspg4",
                      "Acta2",
                      "Adra1a",
                      "Ren1"
                            )

DotPlot(SO3,
        features = markers.to.plot1,
        dot.scale = 8,
        dot.min = 0,
        scale.max = 100,
        scale.min = 0,
        col.min = -2.5,
        col.max = 2.5,
        scale = FALSE,
        group.by = "class.Merge") +
  coord_flip() +
  theme_classic() +
  ggtitle("Populations Markers") +
  theme(axis.line = element_line(size = 1, colour = "black"),
        axis.ticks.x = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.border = element_blank(),
        panel.background = element_blank(),
        text = element_text(size = 20),
        axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5)) +
  xlab(NULL) +
  ylab(NULL)

```

# Save Final Object

```{r echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

saveRDS(SO3, here("Outputs", "GFPcombined.072824.rds"))

#GFPcombined <- readRDS(here("Outputs", "GFPcombined.072824.rds"))

```

# Session Info

```{r sessioninfo, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.align = 'center'}

sessionInfo()

```

